\pkg{Rgbp}: Bayesian Hierarchical Modeling and Frequentist Method Check

### 2013-05-01 jss/article.tex
changes

For example, the shrinkage estimate of the first hospital ($\hat{B}_{1}= 0.834$) was calculated by 336 / (336 + 67), where 67 is its caseload ($n_{1}$), and its posterior mean is $(1-0.834)*0.045 + 0.834 * 0.020=0.024$. As for this hospital, using more information from the prior distribution is a sensible choice because its observed amount of information (67) is far less than the amount of state-level information (336).
\\

@@

It estimated coverage probabilities by simple unbiased estimates (a blue dashed line) and Rao-Blackwellized unbiased estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.951 from the Rao-Blackwellized estimate and 0.95 from the simple unbiased estimate. Considering this result, we finally conclude that the interval estimates from the suggested multilevel modeling on this particular data set has nice repeated sampling property.
\\

@@

The reason we tried true shrinkage values between 0.01 and 0.99 is that no shriankge ($\equiv r / (r + n_{j})=0)$, corresponding to $r=0$, makes the scale of the distribution go to $\infty$ (see (5)), and full shrinkage ($B_{j}=1$) makes $r$ goes to $\infty$, not appropriate for an argument.
\\

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{CodeChunk}
\begin{CodeInput}
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for whole units:

@@

First of all, the independent Binomial distribution given a true batting average of each player (see (7)) seems an appropriate probability distribution describing the uncertainty in this data set. Assuming this distribution, we know that the sample mean (MLE) is an uniformly minimum variance unbiased estimator (UMVUE) with nice asymptotic properties. And it is the simplest one to estimate their true batting averages. But can we say this is the best way to predict their batting averages?
\\

@@

Our hierarchical modeling is a sequel to these researches, which can add one more unique value. Suppose you strongly believe that the outfielder's batting averages should be different from other positions' ones, not giving a credit to the assumption that all players' batting averages shrink towards the same mean. Instead, you want to assume two different second-level hierarchies for outfielders and for other positions, within each of which sharing information and regressing toward the mean (RTTM) occur. This perspective considers true batting averages of outfielders as sampled from the one and those of other positions from the other. Our multilevel modeling provides a way to incorporate such information (covariate) seamlessly into the second-level hierarchy.
\begin{CodeChunk}
\begin{CodeInput}
R> b <- gbp(z, n, x, model = "br")
R> b
\end{CodeInput}
\begin{CodeOutput}
Summary for whole units:

         obs.mean  n   X1 prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.400 45 1.00      0.310     0.715    0.248     0.335    0.429   0.046
2           0.378 45 1.00      0.310     0.715    0.244     0.329    0.420   0.045
3           0.356 45 1.00      0.310     0.715    0.240     0.323    0.411   0.044
4           0.333 45 1.00      0.310     0.715    0.236     0.316    0.403   0.043
5           0.311 45 1.00      0.310     0.715    0.230     0.310    0.396   0.042
6           0.311 45 0.00      0.233     0.715    0.179     0.256    0.341   0.041
7           0.289 45 0.00      0.233     0.715    0.175     0.249    0.331   0.040
8           0.267 45 0.00      0.233     0.715    0.171     0.243    0.323   0.039
9           0.244 45 0.00      0.233     0.715    0.166     0.237    0.315   0.038
10          0.244 45 1.00      0.310     0.715    0.210     0.291    0.379   0.043
11          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
12          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
13          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
14          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
15          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
16          0.200 45 0.00      0.233     0.715    0.155     0.224    0.302   0.038
17          0.178 45 0.00      0.233     0.715    0.148     0.218    0.297   0.038
18          0.156 45 0.00      0.233     0.715    0.140     0.211    0.292   0.039
colMeans    0.265 45 0.44      0.267     0.715    0.191     0.267    0.351   0.041
\end{CodeOutput}
\end{CodeChunk}
Our model reflects on the additional information, \emph{i.e.}, indicator covariate (1 for outfielder and 0 for other positions), estimating two different prior means, 0.31 and 0.23. Also note that shrinkage estimates are the same for all players. It makes sense because shrinkage ($B_{j}\equiv r / (r+n_{j})$) is determined by the relative amount of information between the first-level and the second-level hierarchies. The fact that all players have the same amount of observed information ($n_{j}=45$) and $\hat{r}=\textrm{exp}(4.727)=113$ from below summary tell why the estimated shrinkage is 0.715 (=113 / (113 + 45)).
\\

For reference, we can run the Normal hierarchical model using $y_{j}=z_{j} / n_{j}$ and $V_{j}=\bar{y}(1-\bar{y})/n$, where $\bar{y}=\sum_{j} z_{j} / \sum_{j} n_{j}$, $j=1,\ldots, 18$. As for the point estimates, we calculated squared error losses, $\sum_{j=1}^{18}(p_{j, remaining}-\hat{p}_{j})^{2}$, using the estimates from the Binomial and Normal multilevel models and also using the sample mean (MLE). These values were 0.042, 0.043, and 0.086 each. Note that estimates from the Normal and Binomial models are twice better than MLE in terms of squared error loss. For reference, the shrinkage estimate of the Normal model was 0.708 slightly less than that of the Binomial.
\\

@@

Note that the observed sample mean among outfielders is 0.308 and that among other positions is 0.231. But why are the estimated two prior means from this Binomial model 0.310 and 0.233 each? This can be attributed to the logistic regression which estimates prior mean by a non-linear logit function of $x'\beta$, causing a small bias (see 3.3). The Normal model can avoid this small bias because it uses a linear regression to estimate those prior means.
\\

The 95\% interval plot shows us range of true batting average for each player. Note that observed sample means are all within the interval estimates. And these interval estimates also catched the batting averages over the ramaining season successfully, although the result was left out. Then, how much can we trust these interval estimates? The following method check will answer it.

>>>>

For example, the shrinkage estimate of the first hospital ($\hat{B}_{1}= 0.834$) was calculated by 336 / (336 + 67), where 67 is its caseload ($n_{1}$), and its posterior mean is $(1-0.834)*0.045 + 0.834 * 0.020=0.024$. As for this hospital, using more information from the prior distribution is an appropriate choice because its observed amount of information (67) is far less than the amount of state-level information (336).
\\

@@

It estimated coverage probabilities by simple unbiased estimates (a blue dashed line) and Rao-Blackwellized unbiased estimates (a red line). Both lines are indistingishable from the horizontal line at 0.95 and the estimated overall average coverage rate is 0.950 from the Rao-Blackwellized estimate and 0.951 from the simple unbiased estimate. Considering this result, we finally conclude that the interval estimates from the suggested multilevel modeling on this particular data set has nice repeated sampling property.
\\

@@

%The reason we tried true shrinkage values between 0.01 and 0.99 is that no shriankge ($\equiv r / (r + n_{j})=0)$, corresponding to $r=0$, makes the scale of the distribution go to $\infty$ (see (5)), and full shrinkage ($B_{j}=1$) makes $r$ goes to $\infty$, not appropriate for an argument.

@@

\subsection[Unknown Second-level Mean and No Covariate]{8 Schools: Unknown Second-level Mean and No Covariate}
\begin{CodeChunk}
\begin{CodeInput}
R> g <- gbp(y, se, model = "gr")
R> g
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

@@

First of all, the independent Binomial distribution given a true batting average of each player (see (7)) seems an appropriate probability distribution describing the uncertainty in this data set. And we know that the sample mean (MLE) under this distribution is an uniformly minimum variance unbiased estimator (UMVUE) with nice asymptotic properties. It is also the simplest way to estimate their true batting averages. But can we say this is the best way to predict their batting averages?
\\

@@

Our hierarchical modeling is a sequel to these researches, which can add one more unique value. Suppose you strongly believe that the outfielder's batting averages are different from other positions' ones, not giving a credit to the assumption that all players' batting averages shrink towards the same mean. Instead, you want to assume two different second-level hierarchies for outfielders and for other positions, within each of which sharing information and regressing toward the mean (RTTM) occur. This perspective considers true batting averages of outfielders as sampled from the one and those of other positions from the other. Our multilevel modeling provides a way to incorporate such information (covariate) seamlessly into the second-level hierarchy.
\begin{CodeChunk}
\begin{CodeInput}
R> b <- gbp(z, n, x, model = "br")
R> b
\end{CodeInput}
\begin{CodeOutput}
Summary for each unit (sorted by n):

         obs.mean  n   X1 prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.400 45 1.00      0.310     0.715    0.248     0.335    0.429   0.046
2           0.378 45 1.00      0.310     0.715    0.244     0.329    0.420   0.045
3           0.356 45 1.00      0.310     0.715    0.240     0.323    0.411   0.044
4           0.333 45 1.00      0.310     0.715    0.236     0.316    0.403   0.043
5           0.311 45 1.00      0.310     0.715    0.230     0.310    0.396   0.042
6           0.311 45 0.00      0.233     0.715    0.179     0.256    0.341   0.041
7           0.289 45 0.00      0.233     0.715    0.175     0.249    0.331   0.040
8           0.267 45 0.00      0.233     0.715    0.171     0.243    0.323   0.039
9           0.244 45 0.00      0.233     0.715    0.166     0.237    0.315   0.038
10          0.244 45 1.00      0.310     0.715    0.210     0.291    0.379   0.043
11          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
12          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
13          0.222 45 0.00      0.233     0.715    0.161     0.230    0.308   0.038
14          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
15          0.222 45 1.00      0.310     0.715    0.202     0.285    0.375   0.044
16          0.200 45 0.00      0.233     0.715    0.155     0.224    0.302   0.038
17          0.178 45 0.00      0.233     0.715    0.148     0.218    0.297   0.038
18          0.156 45 0.00      0.233     0.715    0.140     0.211    0.292   0.039
colMeans    0.265 45 0.44      0.267     0.715    0.191     0.267    0.351   0.041
\end{CodeOutput}
\end{CodeChunk}
Our model reflects on the additional information, \emph{i.e.}, indicator covariate (1 for outfielder and 0 for other positions), estimating two different prior means, 0.310 and 0.233. Also note that shrinkage estimates are the same for all players. It makes sense because shrinkage ($B_{j}\equiv r / (r+n_{j})$) is determined by the relative amount of information between the first-level and the second-level hierarchies. The fact that all players have the same amount of observed information ($n_{j}=45$) and $\hat{r}=\textrm{exp}(4.727)=113$ from below summary tell why the estimated shrinkage is 0.715 (=113 / (113 + 45)).
\\

For reference, we can run the Normal hierarchical model using $y_{j}=z_{j} / n_{j}$ and $V_{j}=\bar{y}(1-\bar{y})/n$, where $\bar{y}=\sum_{j} z_{j} / \sum_{j} n_{j}$, $j=1,\ldots, 18$. For comparison, we calculated squared error losses, $\sum_{j=1}^{18}(p_{j, remaining}-\hat{p}_{j})^{2}$, using the estimates from the Binomial and Normal multilevel models and also using the sample mean (MLE). These values were 0.042, 0.043, and 0.086 each. Note that estimates from the Normal and Binomial models are twice better than MLE in terms of squared error loss. 
\\

@@

Note that the observed sample mean among outfielders is 0.308 and that among other positions is 0.231. But why are the estimated two prior means from this Binomial model 0.310 and 0.233 each? This can be attributed to the logistic regression which estimates prior mean by a non-linear logit function of $x'\beta$, causing a small bias (see 3.3). The Normal model can avoid this small bias because it uses a linear regression to estimate these prior means.
\\

The 95\% interval plot shows us range of true batting average for each player, which clarifies the regression toward the mean (RTTM) within two groups. Let's see the 10th, 14th, and 15th players on the graph for example. They are outfielders but their batting averages (sample mean) are far worse than the first five outfielders. RTTM means that it can happen by their bad luck though in the long run their batting averages will come back to their expected ones as outfielders (blue horizontal line). The fact that their sample means are located at the lower part of their 95\% intervals supports this argument. So, their posterior means (red dots) can also be interpreted as their RTTM. 
\\

One more thing to emphasize is that observed sample means are all within the interval estimates. And these interval estimates also catched the batting averages over the ramaining season successfully, although the result was left out. Then, how much can we trust these interval estimates? The following method check will answer it.
